Typed Functional Genetic Programming

### 2013-07-12 tex/thesis.tex
[tex]

@@

\newenvironment{todo}
{ ~\\[0.5em]
  \textbf{TODO}
  \begin{easylist}[itemize]}
{ \end{easylist}
  ~}

@@

Systematic way of solving inhabitation tree corresponds to
generating terms in their long normal forms
in order from smallest to largest in number of symbols and \lheads.
That means minimizing number of and-nodes in and-or-tree view of point on inhabitation tree.
But since inhabitation tree is a special kind of and-or-tree with alternating layers of 
and-nodes (symbols and \lheads) and or-nodes (types) it is also minimizing number of 
size of whole inhabitation/and-or tree.

\begin{todo}
 & Ten poslední odstavec je napsanej divně, přepsat.
   && říct v něm že na to budem teda používat A*. 
\end{todo}

@@

A* is a general informed search algorithm used for finding least-cost 
path from a given initial state to some goal state in a state space.
According to \cite{AIAMA} A* is the most widely known form of best-first search.

As A* traverses the state space, it follows a path of the lowest expected total cost.
It uses a priority queue of alternative unfinished paths.\\

\begin{todo}
& Mužem nás zajímat buď cesta nebo jen cílovej stav, nás bude zajímat
  cílovej stav - proto nebudem zaznamenávat cestu.
& víc to tu rozepsat hlavně zmínit heuristiku 
a že stačí aby byla optimistická atd..
\end{todo}

\Pseudokod{A*( start , isGoal , heuristic )}{\;

	$open \la$ empty priority queue \;
	$closed \la$ empty set \;\;

@@

	\While{$\neg$ $open$.isEmpty() }{\;

		$state \la$ $open$.popWithLowestF()\;\;

@@

		\For{$next \in$ $state$.nexts() }{\;

		    $newG \la$ $state$.G + dist($state$,$next$)\;\;

@@

\begin{todo}
& náš stavovej prostor je strom takže 
  && vime že $next \not\in open \wedge next \not\in closed$
  && a tedy ani nepotřebujeme udržovat množinu closed
& dist(state,next) = 1 
\end{todo}

... thus ...

\Pseudokod{Our-A*( start , isGoal , heuristic )}{

	$open \la$ empty priority queue \;\;

@@

		$state \la$ $open$.popWithLowestF()\;\;

@@

\begin{todo}
& čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
 nódy co už navštívil při přidávání - VŽDYŤ díky tomu žeje to tree-search 
 (vs. graph search) to neni potřeba - pořádně se podívat do AIAMA jakj
 je terminologie a napsat to sem.
\end{todo}

@@

In order to describe our state space to search in \lets introduce
extended version of our tree representation for typed \lterms from \ref{tree-reps}.
This extension consists in adding a new type of leaf node $(\sigma,\Gamma)$ 
standing for unfinished tree of a type $\sigma$ with context $\Gamma$ to use
as set of building blocks
\footnote{\textbf{TODO}: Tady to napsat nějak líp s tim gamma }.\\

@@

For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.

@@

\Tree
   [.\text{$\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$}
	[.\text{$\lambda x_1 \dots x_n$}	
 		\text{$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$}		 				
	]   
   ]\\

\begin{todo}
 & Discuss that it is OK to expand only the DFS-found node. 
The order of unfinished nodes is irrelevant since generation of each branch is 
independent on generating other branches. Mužem na to koukat jako na takovou tu redukci simetrií, najít to v AIAMA a pořádně popsat.

& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
\end{todo}

@@

>>>>

\usepackage{color}

@@

\newenvironment{todo}
{ ~\\[0.5em]
  {\color{red}\textbf{TODO}}
  \begin{easylist}[itemize]}
{ \end{easylist}
  ~}

@@

Systematic way of solving inhabitation tree corresponds to
generating terms in their long normal forms
in order from smallest to largest in number of symbols and \lheads.

We will achieve this by using \textit{A* algorithm}.

%That means minimizing number of and-nodes in 
%and-or-tree view of point on inhabitation tree.
%But since inhabitation tree is a special kind of 
%and-or-tree with alternating layers of 
%and-nodes (symbols and \lheads) and or-nodes 
%(types) it is also minimizing number of 
%size of whole inhabitation/and-or tree.

@@

A* is a general informed search algorithm used for finding least-cost 
path from a given start state to some goal state in a state space.
According to \cite{AIAMA} A* is the most widely known form of best-first search.

By state space is meant oriented weighted graph with states as vertices and edges 
labeled with numbers corresponding to distance or cost. 
Edge from $s_i$ to $s_j$ means that state 
$s_j$ is reachable from state $s_i$ in one step with cost $dist(s_i,s_j)$.
In pseudocode we will refer to set of states reachable from 
state $s$ in one step as to $s.nexts()$.

Typically we use A* algorithm to find path from one specific start state to another 
specific goal state. But sometimes  
we can be interested only in finding a goal state which 
is specified by some $isGoal$ predicate give to the algorithm as
input (and the state space may contain many such goal states). 
This later variant corresponds to our situation, 
so we will continue by describing this variant.

As A* traverses the state space, it follows a path of the lowest 
expected total cost. The expectation is based on \textit{heuristic}
function which for state $s$ predicts distance to the nearest goal state. 
It uses a priority queue of states as its crucial data structure.
In this queue the priority of state $s$ is 
the total expected cost of the path from start state 
going through $s$ and continuing to the nearest goal state. \\

Here follows A* algorithm written in pseudocode:

\Pseudokod{A*( start , isGoal , heuristic )}{
	$open \la$ empty priority queue \;
	$closed \la$ empty set \;\;

@@

	\While{$\neg$ $open$.isEmpty() }{
		$state \la$ $open$.popStateWithLowestF()\;\;

@@

		\For{$next \in$ $state$.nexts() }{		
		    $newG \la$ $state$.G + dist($state$,$next$)\;\;

@@

Now we will modify this general A* algorithm so
it will be more suited for our needs. We want following 
properties:\\

\begin{enum}
& The state space is a tree.
& $dist(s_i,s_j) = 1$
& We do not want to find one, but $n$ goal states. 
\end{enum}~

For state space which is a tree the condition
$(next \not\in open \wedge next \not\in closed)$ 
is always true, since it is possible
to come to a state only from its parent state. 
This fact results in simplification of algorithm's code and behavior;
we may completely omit the $closed$ set and the checks in the for loop.\\

Those changes result in the following code:

\Pseudokod{Our-A*( start , n , isGoal , heuristic )}{

	$open    \la$ empty priority queue \;
	$results \la$ empty set\;\;

@@

		$state \la$ $open$.popStateWithLowestF()\;\;

		\If{ isGoal( $state$ ) }{

			$results$.insert( $state$ )\;

			\If{ $\vert results \vert = n$ }{
				\Return $results$ \;
			}
		}\;

@@

	\Return results
}

The most important part of A* algorithm is the heuristic function.
A heuristic function must be \textit{admissible} in order to make
A* algorithm find the shortest possible path.

A heuristic function $h$ is said to be \textit{admissible} if it satisfies 
following condition for every state $s$:
$$ h(s) \leq h^{*}(s) $$

where $h^{*}(s)$ is the true length of the shortest path from $s$ to 
the nearest goal state.

In other words, a heuristic function is \textit{admissible} if it is
\textit{optimistic}.

@@

In order to describe our state space to search in \lets introduce
extended version of our tree representation for typed \lterms from \ref{tree-reps}.
This extension consists in adding a new kind of leaf node $(\sigma,\Gamma)$ 
standing for unfinished tree of a type $\sigma$ with context $\Gamma$ to use
as set of building blocks
\footnote{\textbf{TODO}: Tady to napsat nějak líp s tim gamma }.\\

@@

For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet used anywhere in the whole tree.
%included in context $\Gamma$ as variable or constant name.

@@

\Tree
   [.\text{$\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$}
	[.\text{$\lambda x_1 \dots x_n$}	
 		\text{$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$}		 				
	]   
   ]\\[3em]

We should discuss that it is correct to have as successors of a state those
created by expansion of the leftmost unfinished leaf (in contrast with
more broad successor set where any unfinished leaf may be expanded).
What we need to show is that no final state will be omitted by the method 
and that there is no shorter path omitted.

One can see that it is correct by considering that 
the order of expanding unfinished leafs is irrelevant 
since expansion of each unfinished leaf is 
independent from expansion of any other unfinished leaf
(up to renaming of variables which is irrelevant).

Every unfinished leaf of a state must be replaced by 
tree containing no unfinished leafs in order to be final state. 
But since expansion of an unfinished leaf has no impact 
(besides new variable names) on expansion of other unfinished
leafs we may choose the order of expansions arbitrarily. \\

We prefer this variant because:
 \begin{enum}
 	& It has smaller or equal number successors for each state, 
 	  i.e. smaller or equal branching factor.
 	& It makes the state space a tree, in contrast with
 	  the naive state space which is not a tree. 
 \end{enum}

\begin{todo}
Mužem na to koukat jako na takovou tu redukci simetrií, 
najít to v AIAMA a pořádně popsat.
\end{todo}

@@

\subsection{Search strategy}

\begin{todo}
 & Internal state
 & Edge selection model

\begin{todo}

& sjednotit terminologii nexts successors atd

& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
